---
title: Grapevine xf Analysis 2
author: Daniel Turek
output: html_document
references:
  - id: tibbits14
    title: Automated factor slice sampling
    author:
    - family: Tibbits
      given: Matthew M
    - family: Groendyke
      given: Chris
    - family: Haran
      given: Murali
    - family: Liechty
      given: John C
    container-title: Journal of Computational and Graphical Statistics
    volume: 23
    issue: 2
    publisher: Taylor & Francis
    page: 543-563
    type: article-journal
    issued:
      year: 2014
---

<!---
danielturek.github.io/public/folder/fileName.html
-->


```{r, include = FALSE}
setwd('~/github/zeilinger_grapevine/')
```

\   

### Updates in Analysis 2

\  

- Vector xf concentration 
  $(\lambda^i_{vector})$ a modelled as a linear function of the source
  plant xf concentration $(\lambda^i_{plant})$.  Further, this linear
  relationship includes an interaction with genotype, so we separately
  model this linear relationship for each source plant genotype.

- Tried introducing a probabilistic model for whether the vectors
  acquire *any* xf bacteria from the source plant, as a function of
  plant genotype, to account for the "patchiness" of the bacteria in
  DSF plants.  However, this introduced a lack of identfyability
  between the probability of *transmitting* bacteria to the vectors (not
  currently in the model) and the probability of *observing* bacteria
  in the vectors (in the model).  So this probabilistic model for
  transmission from source plants to vectors was removed.

- The model for test plant infection was augmented to include a
  dependence on plant genotype.

- Simplified the parameterizations of interaction terms, for better
  interpretability.  The mathematical description of the hierarchical model more clearly 
  presents the biological processes for genotype DSF and WT
  plants.  And more clear descriptions for the interpretation of results.

- Make use of a new joint sampling algorithm avaialble in NIMBLE, the
  automated factor slice sampler [@tibbits14].  Although this increases runtime, it
  is extremely effective at sampling highly correlated groups of parameters.

- Predictive nodes are added to predict test plant infection, for the
  following two cases:
    - genotype = DSF, distance = 0
    - genotype = WT, distance = 0

\   

### Modeling Decisions

\  

##### Vector xf is linear in source plant xf, by genotype (updated v2)

The xf concentration in the vectors
$(\lambda^i_{vector})$ is modeled as a linear function of the xf concentration in the
source plant $(\lambda^i_{source\ plant})$.  However, this linear
relationship is distinct for each genotype.

For genotype DSF plants:
$$\lambda^i_{vector} = \beta_5 + \beta_7 \cdot \lambda^i_{source\
plant}$$

and for genotype DSF plants:
$$\lambda^i_{vector} = \beta_6 + \beta_8 \cdot \lambda^i_{source\
plant}$$

This allows us to study how the linear relationship between source plant xf
concentration and vector xf concentation varies for each genotype.

\   

##### Source plant and vector observation processes 

To model the imperfect detection of xf bacteria in the source plants and 
vectors, I used a fixed (and separate) probability of detection for 
each.  When measuring the xf concentration in each source plant, there's 
a $p_{obs.\ source\ plant}$ probability of a positive detection (e.g.,
a non-zero measurement).

Similarly, when doing the PCR test to 
measure the xf concentration in each vector, there's 
a $p_{obs.\ vector}$ probability of a positive detection.  Each of
these probabilities of detection are unknown model parameters.

To implement these detection processes, additional binary latent variables are
introduced into the model.  The binary-valued $z^i_{source\ plant}$ and $z^i_{vector}$
variables indicate whether a positive detection was made in each source
plant or vector.

\   

##### Test plant infection process (updated v2)

Probabiliy of infection for the test plants is modeled (on the logit
scale) as a linear function of the vector xf concentration, as well as
a constant term for each plant genotype, as:

$$\text{logit}(p^i_{infect}) = \beta_9 \cdot \text{genotype}_{DSF} +
\beta_{10} \cdot \text{genotype}_{WT} + \beta_{11} \cdot  \lambda^i_{vector}$$

This sheds light on differences between the two plant genotypes, although there is still
not significant effect from the vector xf concentration.


\  

##### Plant Random Effects 

The plant random effects $(\alpha_{plant\ ID})$ were removed from the 
model, since there isn't enough data to support them. 
Specifically, with only 97 observations from 51 unique plants,
there aren't enough observations from each plant to identify the random 
effects.  Including them, sampling of the random effects trades off with the 
genotype fixed effects, and the model never converges. 

\   


### Hierarchical Model (updated v2)

\  

$$
\begin{align}
\text{Priors:} \\
\beta_{1,...,11} &\sim Normal(0, 10000) \\
p_{obs.\ source\ plant},\ p_{obs.\ vector} &\sim Uniform(0, 1) \\
\ \\
\text{Biological model (DSF source plants):} \\
\log(\lambda^i_{DSF\ source\ plant}) &= \beta_1 + \beta_3 \cdot distance^i \\
\lambda^i_{vector} &= \beta_5 + \beta_7 \cdot \lambda^i_{DSF\ source\ plant} \\
\ \\
\text{Biological model (WT source plants):} \\
\log(\lambda^i_{WT\ source\ plant}) &= \beta_2 + \beta_4 \cdot distance^i \\
\lambda^i_{vector} &= \beta_6 + \beta_8 \cdot \lambda^i_{WT\ source\ plant} \\
\ \\
\text{Source plant observation model:} \\
z^i_{source\ plant} &\sim Bernoulli(p_{obs.\ source\ plant}) \\
xf^i_{source\ plant} &\sim Poisson(z^i_{source\ plant} \cdot \lambda^i_{source\ plant}) \\
\ \\
\text{Vector observation model:} \\
z^i_{vector} &\sim Bernoulli(p_{obs.\ vector}) \\
xf^i_{vector} &\sim Poisson(z^i_{vector} \cdot \lambda^i_{vector}) \\
\ \\
\text{Test plant infection model:} \\
\text{logit}(p^i_{infect}) &= \beta_9 \cdot \text{genotype}_{DSF} +
\beta_{10} \cdot \text{genotype}_{WT} + \beta_{11} \cdot  
\lambda^i_{vector} \\  
infected^i &\sim Bernoulli(p^i_{infect}) \\
\end{align}
$$

\   

### Data Setup

\  

A few manipulations are done of the input data.

```{r }
dsf <- readRDS('data/dsf_data_for_nimble.rds')   ## original data file
```

We specify covariates:

- $genotype_{dsf}=1$ when genotype = DSF, and 0 otherwise
- $genotype_{wt}=1$ when genotype = WT, and 0 otherwise 

```{r }
constants <- list(
    N = dsf$dsfData$N,
    ##nplant = dsf$dsfData$nplant,
    ##plantID = dsf$dsfData$plantID,
    distance = dsf$dsfData$distance,
    ## in original dataset: genotype 1=DSF, 2=WT
    genotype_dsf = ifelse(dsf$dsfData$genotype==1, 1, 0),   ## indicator for DSF
    genotype_wt  = ifelse(dsf$dsfData$genotype==2, 1, 0)    ## indicator for WT
    ## note: test plants are all WT genotype
)
```

The values of the $xf_{vector}$ data are truncated to be integers, for
compatibility with the Poisson distribution.

```{r }
data <- list(
    xf_source_plant = dsf$dsfData$xf_source_plant,
    xf_vector = floor(dsf$dsfData$xf_vector),   ## change to integers
    infected = dsf$dsfData$infected
)
```

We also create three additional vectors of indicies.  Thes contain the
indicies of the observed (non-NA) data, for each of $xf_{source\
plant}$, $xf_{vector}$, and $infected$.  These will be used in the
model to skip creating posterior predictive nodes, which speed
up the MCMC and overall mixing considerably.

```{r }
## indices of observed values for each set of data:
constants$obs_ind_plant <- which(!is.na(data$xf_source_plant))
constants$obs_ind_vector <- which(!is.na(data$xf_vector))
constants$obs_ind_infected <- which(!is.na(data$infected))

## and the counts of each:
constants$N_obs_plant <- length(constants$obs_ind_plant)
constants$N_obs_vector <- length(constants$obs_ind_vector)
constants$N_obs_infected <- length(constants$obs_ind_infected)
```


```{r include=FALSE, eval=FALSE}
save(dsf, data, constants, file = 'data/temp1.RData')

setwd('~/github/zeilinger_grapevine/')
load('data/temp1.RData')
```


\   

### Model and MCMC 

\  

Use <a href="https://r-nimble.org/" target="_blank">NIMBLE</a> to build the model and corresponding MCMC algorithm.

\   

```{r, eval=FALSE}
library(nimble)

code <- nimbleCode({
    ## priors:
    for(i in 1:11) {
        b[i] ~ dnorm(0, sd=10000)
    }
    p_obs_plant ~ dbeta(1, 1)   ## probability of xf detection in source plant
    p_obs_vector ~ dbeta(1, 1)  ## probability of xf detection in vector
    ## biological model:
    for(i in 1:N) {
        log(lambda_plant[i]) <- b[1]*genotype_dsf[i] + b[2]*genotype_wt[i] + b[3]*distance[i]*genotype_dsf[i] + b[4]*distance[i]*genotype_wt[i]
        lambda_vector[i] <- b[5]*genotype_dsf[i] + b[6]*genotype_wt[i] + b[7]*lambda_plant[i]*genotype_dsf[i] + b[8]*lambda_plant[i]*genotype_wt[i]
    }
    ## only model observed (non-NA) values of xf_source_plant:
    for(iObs in 1:N_obs_plant) {
        z_plant[obs_ind_plant[iObs]] ~ dbern(p_obs_plant)
        xf_source_plant[obs_ind_plant[iObs]] ~ dpois(z_plant[obs_ind_plant[iObs]] * lambda_plant[obs_ind_plant[iObs]])
    }
    ## only model observed (non-NA) values of xf_vector:
    for(iObs in 1:N_obs_vector) {
        z_vector[obs_ind_vector[iObs]] ~ dbern(p_obs_vector)
        xf_vector[obs_ind_vector[iObs]] ~ dpois(z_vector[obs_ind_vector[iObs]] * lambda_vector[obs_ind_vector[iObs]])
    }
    ## only model observed (non-NA) values of infected:
    for(iObs in 1:N_obs_infected) {
        logit(p_infect[obs_ind_infected[iObs]]) <- b[9]*genotype_dsf[obs_ind_infected[iObs]] + b[10]*genotype_wt[obs_ind_infected[iObs]] + b[11]*lambda_vector[obs_ind_infected[iObs]]
        infected[obs_ind_infected[iObs]] ~ dbern(p_infect[obs_ind_infected[iObs]])
    }
    ## genotype DSF predictive nodes:
    log(lambda_plant_dsf) <- b[1]
    lambda_vector_dsf <- b[5] + b[7]*lambda_plant_dsf
    logit(p_infect_dsf) <- b[9] + b[11]*lambda_vector_dsf
    ## genotype WT predictive nodes:
    log(lambda_plant_wt) <- b[2]
    lambda_vector_wt <- b[6] + b[8]*lambda_plant_wt
    logit(p_infect_wt) <- b[10] + b[11]*lambda_vector_wt
})

inits <- list(
    b = c(rep(0,4), 1, 1, 0, 0, rep(0,3)),
    p_obs_plant = 0.5,
    p_obs_vector = 0.5,
    z_plant = rep(1, constants$N),
    z_vector = rep(1, constants$N)
)

Rmodel <- nimbleModel(code, constants, data, inits)

calculate(Rmodel)

conf <- configureMCMC(Rmodel)

## add specialized automated factor slice samplers:
conf$addSampler(c('b[5]','b[7]'), 'AF_slice')
conf$addSampler(c('b[6]','b[8]'), 'AF_slice')
conf$addSampler(c('b[9:11]'),     'AF_slice')

## add monitors for predictive nodes:
conf$addMonitors('p_infect_dsf', 'p_infect_wt')

Rmcmc <- buildMCMC(conf)

system.time(Cmodel <- compileNimble(Rmodel))
```

```
compiling... this may take a minute. Use 'showCompilerOutput = TRUE' to see C++ compiler details.
compilation finished.
   user  system elapsed 
 12.424   0.290  12.786
```

```{r, eval=FALSE}
system.time(Cmcmc <- compileNimble(Rmcmc, project = Rmodel))
```

```
compiling... this may take a minute. Use 'showCompilerOutput = TRUE' to see C++ compiler details.
compilation finished.
   user  system elapsed 
 11.745   0.296  12.121 
```

```{r, eval=FALSE}
set.seed(0)
system.time(samples <- runMCMC(Cmcmc, 50000, nburnin=10000, nchains=2, returnCodaMCMC=TRUE))
```

```
running chain 1...
|-------------|-------------|-------------|-------------|
|-------------------------------------------------------|
running chain 2...
|-------------|-------------|-------------|-------------|
|-------------------------------------------------------|
   user  system elapsed 
 70.132   0.421  71.023 
```

```{r, include=FALSE, eval=FALSE}
save(samples, file='results/samples.RData')
```

```{r, include=FALSE}
setwd('~/github/zeilinger_grapevine/')
load('results/samples.RData')
```

\   

### Convergence 

We assess convergence of two chains, each with 40,000 post burn-in samples.

\   

The Brooks-Gelman-Rubin diagnostic over two chains of indicates convergence.

```{r, warning=FALSE}
library(coda)
```

```{r }
gelman.diag(samples)
```

\   

The effective sample size of the first chain is satisfactory, with a
minimum of 7,000 effectively independent samples from all posterior
dimensions, resulting from the 40,000 post burn-in samples.

```{r }
round(cbind(
    length = apply(samples[[1]], 2, length),
    ESS    = effectiveSize(samples[[1]])
))
```

\   


### Posterior Mean, Median, and 95% Credible Intervals 

\   

```{r}
res <- round(cbind(
    `CI Lower` = apply(samples[[1]], 2, function(x) quantile(x, 0.025)),
    `Mean`         = apply(samples[[1]], 2, mean),
    `Median`       = apply(samples[[1]], 2, median),
    `CI Upper` = apply(samples[[1]], 2, function(x) quantile(x, 0.975))
), 5)

res
```

\   

### Interpretations (updated v2)

\   

$\beta_1$, the constant term in the log() predictor for 
source plant xf concentration in genotype DSF plants, is
essentially `r round(res["b[1]","Mean"],3)`. Thus, the baseline xf
concentration (at distance=0) in genotype DFS plants is $e^{\beta_1}
\approx$ 62,000,000.

$\beta_2$, the constant term in the log() predictor for 
source plant xf concentration in genotype WT plants, is essentially `r round(res["b[2]","Mean"],3)`,
slightly lower than the genotype DSF constant term.  Thus, the baseline xf
concentration (at distance=0) in genotype WT plants is $e^{\beta_2}
\approx$ 44,000,000.

These values of $\beta_1$ and $\beta_2$ imply that
for the same $distance$, genotype WT plants will have less xf
bacteria than genotype DSF plants.  Specifically, lower in the WT
plants by
a factor of $e^{\beta_2-\beta_1} = e^{-0.346}=$ `r round(exp(res["b[2]","Mean"] -
res["b[1]","Mean"]), 3)`.

\   

$\beta_3$, the coefficient of $distance$ in the log() predictor for 
source plant xf concentration in genotype DSF plants, is essentially
`r round(res["b[3]","Mean"],3)`. For every additional unit of
distance, the source plant xf concentration in DSF plants will decrease by a factor
of $e^{\beta_3}=$ `r round(exp(res["b[3]","Mean"]), 3)`.

$\beta_4$, the coefficient of $distance$ in the log() predictor for 
vector xf concentration in genotype WT plants, is essentially
`r round(res["b[4]","Mean"],3)`.  This is significantly closer to 0 than that
of genotype DSF plants, implying that the xf concentration decreases
 more gradually (as a function of distance) in genotype WT plants, as
compared to genotype DSF plants.  Specicially, for every additional unit of
distance, the source plant xf concentration in WT plants will decrease by a factor
of $e^{\beta_4}=$ `r round(exp(res["b[4]","Mean"]), 3)`.

\   

For DSF source plants, $\beta_5$ and $\beta_7$ determine the linear
relationship for how xf bacteria is transmitted to vectors, as
$\lambda_{vector} = \beta_5 + \beta_7 \cdot \lambda_{source\ plant}$.  We are
interested in the $\beta_7$ coefficient, which is approximately equal to 0.00034.  This means that for DSF source
plants, the vector xf concentration is approximately 0.00034 of the source plant xf concentration.

For WT source plants, $\beta_6$ and $\beta_8$ determine the linear
relationship for how xf bacteria is transmitted to vectors, as
$\lambda_{vector} = \beta_6 + \beta_8 \cdot \lambda_{source\ plant}$.  We are
interested in the $\beta_8$ coefficient, which is approximately equal to 0.00085.  This means that for WT source
plants, the vector xf concentration is approximately 0.00085 of the
source plant xf concentration.  This is a much higher transmission
rate from (source plants to vectors) from WT plants relative to DSF
plants.

\   

$\beta_{9}$ gives the baseline probability of infection (on the logit
scale) for genotype DSF plants.  The 95% CI for $\beta_{9}$ is
(`r round(res["b[9]","CI Lower"], 2)`,
`r round(res["b[9]","CI Upper"], 2)`).  Taking the inverse-logit
transformation, this gives a 95% CI for the baseline probability of infection for DSF
source plants as (`r round(1/(1+exp(-res["b[9]","CI Lower"])), 2)`,
`r round(1/(1+exp(-res["b[9]","CI Upper"])), 2)`).  This
does not take xf concentration in the vector into account.

$\beta_{10}$ gives the baseline probability of infection (on the logit
scale) for genotype WT plants.  The 95% CI for $\beta_{10}$ is
(`r round(res["b[10]","CI Lower"], 2)`,
`r round(res["b[10]","CI Upper"], 2)`).  Taking the inverse-logit
transformation, this gives a 95% CI for the baseline probability of infection for WT
source plants as (`r round(1/(1+exp(-res["b[10]","CI Lower"])), 2)`,
`r round(1/(1+exp(-res["b[10]","CI Upper"])), 2)`).  This
does not take xf concentration in the vector into account.

\   

 $\beta_{11}$, the coefficient for vector xf level in the logit() predictor for test plant
infection, is not significantly different from 0, with 95% CI
(-0.00015, 0.00002).

\   

$p_{obs.\ source\ plant}$, the probability of xf detection in source
plants, is bounded by the 95% CI (`r round(res["p_obs_plant","CI Lower"], 2)`,
`r round(res["p_obs_plant","CI Upper"], 2)`).


$p_{obs.\ vector}$, the probability of xf detection in vectors,
is bounded by the 95% CI (`r round(res["p_obs_vector","CI Lower"], 2)`,
`r round(res["p_obs_vector","CI Upper"], 2)`).

\   


The predictive nodes give the best idea of the probability of
infection, from either DSF or WT source plants, using distance = 0.
These estimates take all model parameters and biological processes into account.

- Probability of infection for genotype DSF is in the 95% CI: (`r round(res["p_infect_dsf","CI Lower"], 2)`,
`r round(res["p_infect_dsf","CI Upper"], 2)`), with median = `r round(res["p_infect_dsf","Median"], 2)`.
- Probability of infection for genotype WT is in the 95% CI: (`r round(res["p_infect_wt","CI Lower"], 2)`,
`r round(res["p_infect_wt","CI Upper"], 2)`), with median = `r round(res["p_infect_wt","Median"], 2)`.

\   




### Posterior Density Plots

\   

Overlay the predictive distributions for test plant infection:


```{r fig.height=2}
samplesPlot(samples[[1]], c('p_infect_dsf','p_infect_wt'), height=2)
```

\  

```{r, eval = FALSE}
plot(samples[[1]], ask = FALSE)
```

```{r, echo = FALSE}
for(i in 1:5)
    plot(coda::as.mcmc(samples[[1]][, (3*i-2):(3*i)]), ask = FALSE)
```



### References

\   


```{r eval=FALSE, include=FALSE}
rng <- range(samples[[1]][, Rmodel$expandNodeNames('lambda_vector')])
par(mfrow=c(2,1)), mar=c(1,1,1,1))
ymax <- 0.03
makePolygons <- function(ind, data, samples) {
    for(i in ind) {
        dens <- density(samples[[1]][, paste0('lambda_vector[',i,']')])
        color <- if(data$infected[i]) 'red' else 'black'
        polygon(x=dens$x, y=dens$y, border=color)
    }
}
plot(0,0, xlim=rng, ylim=c(0,ymax), main='Genotype DSF', xlab='lambda_vector', ylab='posterior density')
ind <- which(!is.na(data$infected) & constants$genotype_dsf)
makePolygons(ind, data, samples)
plot(0,0, xlim=rng, ylim=c(0,ymax), main='Genotype WT',  xlab='lambda_vector', ylab='posterior density')
ind <- which(!is.na(data$infected) & constants$genotype_wt)
makePolygons(ind, data, samples)




rng <- range(c(
    as.numeric(samples[[1]][, Rmodel$expandNodeNames('lambda_vector')]),
    as.numeric(na.omit(data$xf_vector))
))
par(mfrow=c(2,1)), mar=c(1,1,1,1))
ymax <- 0.03
pointsY <- 0
pch <- 19
makePolygons <- function(ind, data, samples) {
    for(i in ind) {
        dens <- density(samples[[1]][, paste0('lambda_vector[',i,']')])
        color <- if(data$infected[i]) 'red' else 'black'
        polygon(x=dens$x, y=dens$y, border=color)
    }
}
plot(0,0, xlim=rng, ylim=c(0,ymax), main='Genotype DSF', xlab='lambda_vector', ylab='posterior density')
ind <- which(!is.na(data$infected) & constants$genotype_dsf)
makePolygons(ind, data, samples)
ind <- which(!is.na(data$xf_vector) & !is.na(data$infected) & constants$genotype_dsf)
points(x=data$xf_vector[ind], y=rep(pointsY,length(ind)), pch=pch, col = ifelse(data$infected[ind], 'red', 'black'))


plot(0,0, xlim=rng, ylim=c(0,ymax), main='Genotype WT',  xlab='lambda_vector', ylab='posterior density')
ind <- which(!is.na(data$infected) & constants$genotype_wt)
makePolygons(ind, data, samples)
ind <- which(!is.na(data$xf_vector) & !is.na(data$infected) & constants$genotype_wt)
points(x=data$xf_vector[ind], y=rep(pointsY,length(ind)), pch=pch, col = ifelse(data$infected[ind], 'red', 'black'))
```





```{r, include=FALSE, eval=FALSE}
library(coda)
apply(samples, 2, length)
apply(samples, 2, mean)
apply(samples, 2, effectiveSize)

cor(samples)   ## b[5] and b[6] correlated

burn <- 0
burn <- 1000
burn <- 5000
burn <- 50000
samplesPlot(samples)
samplesPlot(samples, 'b[1]', burnin=burn)   ## slow mixing b[1]
samplesPlot(samples, 'b[2]', burnin=burn)   ## slow mixing b[2]
samplesPlot(samples, 'b[3]', burnin=burn)
samplesPlot(samples, 'b[4]', burnin=burn)
samplesPlot(samples, 'b[5]', burnin=burn)
samplesPlot(samples, 'b[6]', burnin=burn)
samplesPlot(samples, 'k', burnin=burn)      ## slow mixing k ??
##samplesPlot(samples, 'k2', burnin=burn)      ## slow mixing k ??
samplesPlot(samples, 'p_obs_plant', burnin=burn)
samplesPlot(samples, 'p_obs_vector', burnin=burn)
samplesPlot(samples, 'sigma_alpha', burnin=burn)


codamcmc <- runMCMC(Cmcmc, 50000, nburnin=10000, nchains=3, returnCodaMCMC=TRUE)
coda::gelman.diag(codamcmc, autoburnin = FALSE)




colnames(samples)
samplesPlot(samples, 1)  ## slow mixing b[1]
samplesPlot(samples, 2)  ## slow mixing b[2] 
samplesPlot(samples, 3)
samplesPlot(samples, 4)
samplesPlot(samples, 5)
samplesPlot(samples, 6)
samplesPlot(samples, 7)  ## slow mixing k
samplesPlot(samples, 8)
samplesPlot(samples, 9)
samplesPlot(samples, 10)



##code <- nimbleCode({
##  mu_alpha ~ dnorm(0, 0.001)
##  sigma_alpha ~ dunif(0, 1000)
##  for(i in 1:nplant) { 
##    alpha[i] ~ dnorm(mu_alpha, sd = sigma_alpha)  ## site random effect
##  }
##  for(i in 1:7) {
##    beta[i] ~ dnorm(0, 0.001)
##  }
##  for(i in 1:2) {
##    betagenotype[i] ~ dnorm(0, 0.001)
##  }
##  for(i in 1:N) {
##    # Source plant sub-model
##    log(lambda_plant[i]) <- alpha[plantID[i]] + betagenotype[genotype[i]] + beta[1]*distance[i]
##    xf_source_plant[i] ~ dpois(lambda_plant[i])
##    # Vector sub-model with detection probabilty
##    log(lambda_vector[i]) <- beta[2] + beta[3]*lamba_plant[i] # Biological process
##    logit(p_vector[i]) <- beta[4] + beta[5]*p_trans[i] # Detection probability
##    N_vector[i] ~ dpois(lambda_vector[i]) # True Xf population in vectors
##    xf_vector[i] ~ dbin(p_vector[i], N_vector[i]) # Data on Xf population in vectors
##    # Transmission probability sub-model
##    logit(p_trans[i]) <- beta[6] + beta[7]*N_vector[i]
##    infected[i] ~ dbern(p_trans[i])
##  }
##})
## 
##constants <- with(nimbleTransData,
##                  list(N=N, 
##                       nplant=nplant, 
##                       genotype = genotype, 
##                       distance = distance,  
##                       #genotype_distance = genotype_distance, 
##                       # xf_source_plant = xf_source_plant,
##                       # xf_vector = xf_vector,
##                       # infected = infected,
##                       plantID = plantID))
## 
##data <- with(nimbleTransData, 
##             list(xf_source_plant = xf_source_plant,
##                  xf_vector = xf_vector,
##                  infected=infected))
## 
##inits <- list(mu_alpha=0, sigma_alpha=1, alpha=rep(0,nimbleTransData$nplant), beta=rep(0,7), betagenotype=rep(0,2))
## 
##modelInfo <- list(code=code, constants=constants, data=data, inits=inits, name='month')
## 
## 
###### Set up model and samplers
##Rmodel <- nimbleModel(modelInfo$code,
##                      modelInfo$constants,
##                      modelInfo$data,
##                      modelInfo$inits)
## 
##Cmodel <- compileNimble(Rmodel)
## 
##spec <- configureMCMC(Rmodel)
## 
###### Best configuration of samplers for random effect occupancy model
##spec$removeSamplers('beta[1:9]')
##spec$addSampler('beta[1:3]', 'RW_block') # detection sub-model sampler
##spec$addSampler('beta[4:9]', 'RW_block') # occupancy sub-model sampler
##spec$removeSamplers('sigma_alpha')
##spec$addSampler('sigma_alpha', 'RW_log_shift', list(shiftNodes='alpha')) # random effect sampler
##spec$getSamplers() # Check samplers
##spec$addMonitors(c('p_occ')) # add a monitor to get p_occ in output
###spec$addMonitors(c('p_obs')) # add a monitor to get p_obs in output
## 
###### Compile MCMC in R and C++
##Rmcmc <- buildMCMC(spec)
##Cmcmc <- compileNimble(Rmcmc, project = Rmodel)
## 
###### Run MCMC with 150,000 iterations and 50,000 burn-in
##niter <- 1500
##burnin <- 500
## 
##ti <- Sys.time()
##samplesList <- lapply(1, mcmcClusterFunction)
##tf <- Sys.time()
## 
### The time it took to run MCMC
##tf-ti
## 
##save(samplesList, file = 'output/MCMC_list_climate_transmission.RData')


if(FALSE) {
    ## pcr data, possibly to use later
    pcr <- readRDS('data/xf_vector_population_pcr_data.rds')  ## new file with the replicated bacteria counts for vectors
    str(pcr)
    head(pcr, 15)
    dim(pcr)
    ## We use quantitative PCR (qPCR), which means we are estimating the number of copies of a gene fragment specific to the bacteria in the tissues of the insect vector. We extract all the DNA from each insect, then take two small samples of the extract and run qPCR in duplicate. The xf_vector numbers that I sent you is actually the mean of the two duplicate qPCR runs for each sample. I've attached the data showing the duplicates. The "sample" column is the ID of the insect vector, "rep1" and "rep2" are the replicates of the counts of bacteria, and "mean_xf_vector" is the mean of rep1 and rep2, and also the data that is in the "xf_vector" variable I sent you earlier.  As you can see, replicates for each vector are either both 0 or both >0; there are only 4 vectors where one replicate is 0 and the other is >0. Additionally, when both replicates are >0, there is some variation in the estimated count between the two replicates. 
}
##


##ind <- 43
##ind <- 44
##ind <- 56
##ind <- 78
##ind <- 96
##ind <- 97
## 
##nodes <- c(
##    paste0('b[',6,']'),
##    ##'sigma_alpha', 'mu_alpha',
##    'p_obs_plant', 'p_obs_vector',
##    ##'k',
##    paste0('z_plant[',ind,']'),
##    paste0('z_vector[',ind,']'),
##    paste0('lambda_plant[',ind,']'),
##    paste0('lambda_vector[',ind,']'),
##    paste0('p_infect[',ind,']'),
##    paste0('xf_source_plant[',ind,']'),
##    paste0('xf_vector[',ind,']'),
##    paste0('infected[',ind,']')
##)
##i <- 5000:5010
##cbind(samples[i, nodes],
##      dat_source=data$xf_source_plant[ind],
##      dat_vector=data$xf_vector[ind],
##      dat_infect=data$infected[ind])
```

